An ultra-low noise, high-voltage piezo driver

### 2016-01-26 piezo.tex
Reworked intro to circuit description
* Repartitioned "intro" and flyback regulator section; other mods

Traditionally, laboratory electronics capable of supplying high voltages fall under one of two architectural umbrellas: switching converters, and ``linear'' amplifiers.
DC-DC converters are efficient and can work at very high voltages, but suffer from switching noise and limited control bandwidths.
Linear-type devices are typically constructed from a high-voltage operational amplifier (op-amp), powered either from a high voltage linear regulator or more typically from a secondary switching converter.
While the op-amp provides \SI{100}{\decibel} or more of power-supply noise rejection, a high-voltage op-amp requires substantially more power than an equivalent switching circuit and is thus more cumbersome to deploy in the laboratory.

Here, we present an architecture which combines the best of both worlds: a high-voltage piezoelectric driver with excellent noise characteristics that can be powered from a standard low-current, low-voltage supply.
The DC operating point is controlled digitally, and an analog input provides a DC-coupled, unity-gain feedback path ($\pm\SI{10}{\volt}$) to the output.
The following design is versatile and easy to deploy throughout the lab in a variety of high-voltage applications.

@@

\begin{figure*}[t]
\includegraphics[width=\textwidth]{fig/DRV2700.ai}
\caption{Schematic of the high voltage stabilization.
The voltage HV is generated using a Texas Instruments DRV2700 high voltage driver in flyback configuration (see Fig.~\ref{Fig:DRV2700}).
A fast, very high slew-rate op-amp senses the output voltage across $R_1$ and $R_2$, and servos it by modulating the node at ``HV floating gnd''.
The $V_{\text{DC}}$ gain is set by $\left(1+R_1/R_2\right)$, while the modulation gain is set by $-R_{\text{mod}}/R_1$.
The capacitor linking the floating ground node to the output allows the op-amp to remove residual switching noise and stabilize the DC output according to the transfer function given in Eq.~(\ref{Eq:PiezoTransfer}). -- Add back in second set of shunt caps!\label{Fig:PiezoCircuit}}
\end{figure*}

The design is based on the newly-available Texas Instruments DRV2700 piezo driver\footnote{The identification of commercial products is for information only and does not imply recommendation or endorsement by the National Institute of Standards and Technology.}, however it is easily modified to work with any galvanically isolated DC-DC converter.
This single-chip integrated circuit (IC) can be operated as a boost converter to drive an on-chip differential amplifier up to \SI{100}{\volt}, or as a flyback converter up to \SI{1}{\kilo\volt}.
In flyback configuration, the internal-boost switch of the DRV2700 drives a step-up transformer.
When the switch closes, current begins to flow through the primary coil of the transformer and induces a corresponding voltage across the secondary coil.
In this state, the output diode is reverse-biased, and the capacitor ($\text{C}_{\text{HV}}$ in Fig.~\ref{Fig:PiezoCircuit}) holds its charge.
When the switch opens, the voltage across the secondary coil is inverted, putting the diode into conduction and charging the capacitor.
By changing the rate at which the switch is engaged, the DRV2700 is able to regulate a particular voltage at the output.

In isolation, the DRV2700 is not suited for low-noise laboratory instrumentation. 
The output ripple, even after heavy filtering, can be as high as a few volts, and standard filtering techniques to reduce this noise simultaneously reduce the modulation bandwidth of the flyback regulator.
Even if this were not the case, it remains unlikely that the RMS noise could be brought below $\SI{1}{\milli\volt}$, which is the operating regime we require for certain low-noise applications in the laboratory.
Despite this drawback, the flyback converter requires very little current to operate and leverages a compact IC to generate high voltages without the need of an auxiliary supply.

The architecture presented in this paper uses an active feedback element (U2, described in Sec.~\ref{Sec:LowNoiseStabilization}) to remove switching noise and residual ripple at the high-voltage output of the flyback regulator.
Additionally, U2 provides a DC-coupled, unity-gain feedback path to the high voltage node, which can be used for closed-loop servo applications and high-bandwidth modulation.
This design shows how researchers might leverage the versatility of galvanically isolated switching regulators (such as the DRV2700) in a low-noise laboratory environment. [versatile and easy to deploy in the lab.]

In the next few sections, we provide details on each sub-block of the circuit and draw attention to important design details.
The full schematic, including Gerbers, bill of materials, and layout design files, can be found on GitHub~\footnote{\protect\url{https://github.com/JQIamo/piezo-compact-high-voltage-driver}}.

\subsection{DRV2700 flyback regulator}
\label{Sec:DRV2700}

We based the flyback regulator design off the suggested schematic in the DRV2700 datasheet and evaluation module application note.
At a basic level, the DRV2700 flyback circuit requires a transformer (ATB3225; 1:10 step-up winding, \SI{7}{\micro\henry} inductance, rated for \SI{0.6}{\ampere}) connected to an output diode and capacitor, and a ``sense'' voltage supplied to the feedback (FB) node of the IC.
Internally, the DRV2700 boost controller tries to drive the FB node to \SI{1.3}{\volt}.
The op-amp U1 senses the voltage at node HV and $\text{GND}_\text{HV}$, and adjusts its output such that
\begin{align}
\label{Eq:U1Output}
\text{HV} &= G\cdot V_{\text{dc}} + \text{GND}_{\text{HV}}\,,
\end{align}
where the gain $G$ is set by the resistor ratio $R_3/R_4 \equiv R_5/R_6$.
The capacitors $C_3$ and $C_4$ are chosen such that $C_3 = \SI{22}{\pico\farad}$ and 
\begin{align}
\frac{C_4}{C_3} &= \frac{R_3}{R_4~||~R_5}\,,
\end{align}
as suggested by the datasheet.
This pseudo-differential configuration ensures that, in the equilibrium given by Eq.~(\ref{Eq:U1Output}), the output of U1 is driven to \SI{0}{\volt}; the resistive divider $R_9$ and $R_{10}$ is then chosen such that $R_9/(R_9~||~R_{10}) = \SI{1.3}{\volt}/\SI{5}{\volt} \approx \num{0.26}$.

The galvanically isolated output of the flyback regulator is passed through a four-pole, low-pass RC filter (corner frequency $f_c \approx \SI{3}{\kilo\hertz}$). 
Additional capacitors on both the HV and $\text{GND}_{\text{HV}}$ resistor networks shunt high frequency noise to ground. 
The DRV2700 IC, combined with the feedback provided by U1, regulates the voltage difference $(\text{HV}-\text{GND}_{\text{HV}}) = G\cdot V_{\text{dc}}$. The gain $G = 50$ for the design implemented in our lab, allowing a \SI{5}{\volt} DAC to span \SI{0}{\volt} -- \SI{250}{\volt}.

@@

The low-noise stabilization circuit is crucial to the performance of the design, as it is responsible for removing noise at the output of the flyback converter.
To accomplish this task, a high slew-rate op-amp (Texas Instruments LM7171, \SI{4100}{\volt\per\micro\second}) drives the galvanically isolated ground node of the flyback converter (see U2 in Fig.~\ref{Fig:PiezoCircuit}).
This op-amp enforces the HV output voltage
\begin{align}
\text{V}_\text{HV} &= \left(1 + \frac{R_1}{R_2~||~R_\text{mod}}\right) \text{V}_\text{DC} -
\left(\frac{R_\text{mod}}{R_1}\right) \text{V}_\text{mod}
\end{align}
We choose $R_1, R_\text{mod} = \SI{1}{\mega\ohm}$, $R_2 = \SI{20.5}{\kilo\ohm}$ such that the DC gain is $\approx 50$ and the modulation gain is unity.
Depending on the application, however, other gain configurations would work equally well provided the non-inverting gain of U2 closely matches the gain of the flyback regulator since they derive from the same control voltage.
Ultimately, the op-amp U2 sets the operating voltage of the output, while the flyback converter regulates the high voltage difference between $\text{GND}_\text{HV}$ and the output.
Because the output of the flyback regulator is galvanically isolated, U2 is free to control the $\text{GND}_\text{HV}$ node at ``low'' voltages, which are transferred directly to the high voltage output.

@@

The DC setpoint is controlled by a low-noise digital-to-analog (DAC) converter.
This has several advantages: digital control enhances repeatability, and makes it easy to interface with a wide variety of control electronics to, eg, implement a slow infinite integrator [expand on this?]. 
A wide array of high-resolution DACs are available with very good noise characteristics.
We initially chose to use an AD5563R, but found that our performance was ultimately limited by $1/f$ noise in the DAC itself.
The final design uses an XXX, with XXX external reference, such that the noise contribution from the DAC is comparable to the noise contribution from the LM7171 (see Sec.~\ref{Sec:NoiseAnalysis}).

>>>>

%Here, we present an architecture which combines the best of both worlds: a high-voltage piezoelectric driver with excellent noise characteristics that can be powered from a standard low-current, low-voltage supply.
%The DC operating point is controlled digitally, and an analog input provides a DC-coupled, unity-gain feedback path ($\pm\SI{10}{\volt}$) to the output.
%The following design is versatile and easy to deploy throughout the lab in a variety of high-voltage applications.

@@

\begin{figure*}[t]
\includegraphics[width=\textwidth]{fig/DRV2700.ai}
\caption{Schematic of the high voltage stabilization.
The voltage HV is generated using a Texas Instruments DRV2700 high voltage driver in flyback configuration.
A fast, very high slew-rate op-amp senses the output voltage across $R_1$ and $R_2$, and servos it by modulating the node at ``HV floating gnd''.
The $V_{\text{DC}}$ gain is set by $\left(1+R_1/R_2\right)$, while the modulation gain is set by $-R_{\text{mod}}/R_1$.
The capacitor linking the floating ground node to the output allows the op-amp to remove residual switching noise and stabilize the DC output according to the transfer function given in Eq.~(\ref{Eq:PiezoTransfer}). -- Add back in second set of shunt caps!\label{Fig:PiezoCircuit}}
\end{figure*}

Traditionally, laboratory electronics capable of supplying high voltages fall under one of two architectural umbrellas: DC-DC switching converters, and ``linear'' amplifiers.
DC-DC converters are efficient and can work at very high voltages, but suffer from switching noise and limited control bandwidths.
Linear-type devices are typically constructed from a high-voltage operational amplifier (op-amp), powered either from a high voltage linear regulator or more typically from a secondary switching converter.
While the op-amp provides \SI{100}{\decibel} or more of power-supply noise rejection, high-voltage op-amps requires substantially more power than an equivalent switching circuit and are more cumbersome to deploy in the laboratory.

Here, we present a hybrid architecture (see Fig.~\ref{Fig:PiezoCircuit} for details).
The high voltage is generated by a galvanically isolated switching converter (described in Sec.~\ref{Sec:DRV2700}), while a low-noise, high slew-rate op-amp removes the switching noise by modulating the \textbf{OUT--} node of the DC-DC converter (Sec.~\ref{Sec:LowNoiseStabilization}).
This op-amp additionally provides a low-gain, high-bandwidth modulation input for closed-loop feedback applications.

The design principles discussed below show how researchers might leverage the versatility of galvanically isolated switching regulators in a low-noise laboratory environment; straightforward modifications to the schematic make it possible to tailor the high voltage driver to a specific application.
The entire electronics package fits comfortably into a 12HP eurocard rack module, and draws less than \SI{50}{\milli\ampere} at \SI{15}{\volt} [get real numbers].
The full schematic, including Gerbers, a bill of materials, and layout design files, can be found on GitHub~\footnote{\protect\url{https://github.com/JQIamo/piezo-compact-high-voltage-driver}}.

\subsection{Flyback regulator}
\label{Sec:DRV2700}

The high voltage DC-DC converter used here is based on the newly-available Texas Instruments DRV2700 piezo driver\footnote{The identification of commercial products is for information only and does not imply recommendation or endorsement by the National Institute of Standards and Technology.}.
This single-chip integrated circuit (IC) can be operated as a boost converter to drive an on-chip differential amplifier up to \SI{100}{\volt}, or as a flyback converter up to \SI{1}{\kilo\volt}.
In flyback configuration, the internal-boost switch of the DRV2700 drives a step-up transformer.
When the switch closes, current begins to flow through the primary coil of the transformer and induces a corresponding voltage across the secondary coil.
In this state, the output diode is reverse-biased, and the capacitor ($\text{C}_{\text{HV}}$ in Fig.~\ref{Fig:PiezoCircuit}) holds its charge.
When the switch opens, the voltage across the secondary coil is inverted, putting the diode into conduction and charging the capacitor.
By changing the rate at which the switch is engaged, the DRV2700 is able to regulate a particular voltage across the galvanically isolated output (nodes $\text{GND}_\text{HV}$ and HV in Fig.~\ref{Fig:PiezoCircuit}).

The transformer (ATB3225, 1:10 step-up winding), diode, and RC feedback network are based on values suggested in the DRV2700 datasheet~\cite{DRV2700Datasheet,DRV2700EVMUserGuide}.
Internally, the DRV2700 boost controller tries to drive the FB node to \SI{1.3}{\volt}.
The op-amp U1 senses the voltage at node HV and $\text{GND}_\text{HV}$, and adjusts its output such that
\begin{align}
\label{Eq:U1Output}
\text{HV} - \text{GND}_{\text{HV}} &= G\cdot V_{\text{dc}}\,,
\end{align}
where the gain $G$ is set by the resistor ratio $R_3/R_4 \equiv R_5/R_6$.
The capacitors $C_3$ and $C_4$ are chosen such that $C_3 = \SI{22}{\pico\farad}$ and 
\begin{align}
\frac{C_4}{C_3} &= \frac{R_3}{R_4~||~R_5}\,,
\end{align}
as suggested by the datasheet, and $R_6 = R_7 = R_8$.
In our lab, we implement a gain $G\approx 50$ ($R_3 = R_5 = \SI{499}{\kilo\ohm}$; $R_4 = R_{6-8} = \SI{10}{\kilo\ohm}$) such that a \SI{5}{\volt} digital-to-analog (DAC) converter at $V_\text{DC}$ can span \SI{0}{\volt} -- \SI{250}{\volt}.

This pseudo-differential configuration ensures that, in the equilibrium given by Eq.~(\ref{Eq:U1Output}), the output of U1 is driven to \SI{0}{\volt}; the resistive divider $R_9$ and $R_{10}$ is then chosen such that $R_{10}/(R_9~||~R_{10}) = \SI{1.3}{\volt}/\SI{5}{\volt} \approx \num{0.26}$.

The output of the flyback regulator is passed through a four-pole, low-pass RC filter.
The corner frequency $f_c \approx \SI{3}{\kilo\hertz}$ is chosen to be high enough that a slow ($\approx \SI{10}{\hertz}$) rail-to-rail triangle ramp can be applied at $V_\text{DC}$ (for, e.g., sweeping over a spectroscopy resonance), but low enough that the $\approx \SI{100}{\kilo\hertz}$ switching noise is attenuated.
Additional capacitors on both the HV and $\text{GND}_{\text{HV}}$ resistor networks shunt high frequency noise to ground.

@@

The low-noise stabilization circuit is crucial to the performance of the design, as it is responsible for removing noise at the output of the flyback converter.
To accomplish this task, a high slew-rate op-amp (Texas Instruments LM7171, \SI{4100}{\volt\per\micro\second}) drives the galvanically isolated ground node of the flyback converter (see U2 in Fig.~\ref{Fig:PiezoCircuit}).
This op-amp enforces the HV output voltage
\begin{align}
\text{V}_\text{HV} &= \left(1 + \frac{R_1}{R_2~||~R_\text{mod}}\right) \text{V}_\text{DC} -
\left(\frac{R_\text{mod}}{R_1}\right) \text{V}_\text{mod}
\end{align}
We choose $R_1, R_\text{mod} = \SI{1}{\mega\ohm}$, $R_2 = \SI{20.5}{\kilo\ohm}$ such that the DC gain is $\approx 50$ and the modulation gain is unity.
Depending on the application, however, other gain configurations would work equally well provided the non-inverting gain of U2 closely matches the gain of the flyback regulator since they derive from the same control voltage.
Ultimately, the op-amp U2 sets the operating voltage of the output referenced to true circuit ground, while the flyback converter regulates the high voltage relative to the pseudoground node $\text{GND}_\text{HV}$.
Because the output of the flyback regulator is galvanically isolated, U2 is free to control the $\text{GND}_\text{HV}$ node at ``low'' voltages, which are transferred directly to the high voltage output.

@@

The DC setpoint is controlled by a low-noise digital-to-analog (DAC) converter.
This has several advantages: digital control enhances setpoint repeatability, and makes it easy to interface with a wide variety of control electronics to, eg, implement a slow infinite integrator [expand on this?]. 
A wide array of high-resolution DACs are available with very good noise characteristics.
We initially chose to use an AD5563R, but found that our performance was ultimately limited by $1/f$ noise in the DAC itself.
The final design uses an XXX, with XXX external reference, such that the noise contribution from the DAC is comparable to the noise contribution from the LM7171 (see Sec.~\ref{Sec:NoiseAnalysis}).
